---
title: "Holobiont project Random Forest"
author: "Arthur & Leticia"
date: "12/10/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Random Forest Evaluation
This document contains the code witcht performs the random Forest method with the objective to extract from the data:

- The individuals that are most important to classify the variables in this study.  
- etc.  

You can find more about the method in the [Breiman's paper]("https://link.springer.com/article/10.1023/A:1010933404324")

## Libraries we used
```{r libraries, warning=FALSE,message=FALSE}
library(tidyverse)
library(randomForest)
library(scales)
#library(cowplot)
#library(extrafont)
library(patchwork)
```
## Importing and tidying the data
```{r reading data}
data <- read.csv("input/detailed_abundance_phyla_157_corais_matrix_porcentagem_prot_divided_17abr2019.csv") %>%
  select(-1)

levels(data$health_status)[levels(data$health_status) == "iminent bleaching"] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "signs of bleaching"] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "bleaching event"] <- "bleached" 
levels(data$health_status)[levels(data$health_status) == "bleaching event "] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "Cyanobacterial patches"] <- "Black Band Disease"

data_num <- data %>% select_if(is.numeric)

abundantes <- data_num %>% gather() %>% group_by(key) %>% 
  summarise(mean_abun = mean(value)) %>% 
  filter(mean_abun > 1/100) %>% pull(key)


data <- data %>% 
  mutate(
    seq_method = factor(
      case_when(
        .$seq_method %in% c("Illumina Genome Analyzer IIx", "Illumina HiSeq 1000",
                            "Illumina HiSeq 1000 WGS", "Illumina HiSeq 2000",
                            "Illumina HiSeq 2500", "Illumina HiSeq 2500 WGS",
                            "WGS illumina") ~ "Illumina",
        TRUE ~ "454"
      )
    )
  )

data <- data %>% filter(genus != "Fungia",
                        genus != "Pseudodiploria", 
                        !(.$genus == "Porites" & .$authors == "Thurber et al. 2009"),
                        !(.$genus == "Acropora" & .$region == "mesocosmos" & .$authors == "not published"),
                        !(.$genus == "Pocillopora" & .$region == "mesocosmos" & .$authors =="not published"),
                        !(.$genus == "Seriatopora" &.$region == "mesocosmos" & .$authors =="not published"),
                        !(.$id == "mgm4694757_prinseq_good_Crf1")) %>% 
  mutate(genus = factor(genus),
         authors = factor(authors),
         growth_form = factor(growth_form))
```

## Runinng the model
```{r RF,cache=FALSE}
set.seed(2019)

community_structure <- randomForest(importance = TRUE, ntree = 50000,
                                    x = select(data, genus, seq_method, authors, region, growth_form,
                                               health_status,             
                                               Unclassified.class.in.Proteobacteria:Rhodothermaeota)
                                    )
genus_supervised <- randomForest(genus ~ .,
                                 importance = TRUE, ntree = 50000,
                                 select(data, genus, seq_method, authors, region, growth_form,
                                        health_status,             
                                        Unclassified.class.in.Proteobacteria:Rhodothermaeota)
                                 )

full_join(
  community_structure %>% importance() %>% as.data.frame() %>% rownames_to_column(var = "variables") %>% 
    select(MeanDecreaseAccuracy, MeanDecreaseGini,variables) %>% 
    mutate(rf_type = "community_structure") %>% 
    filter(variables != "genus") %>% 
    rename(mda_community = MeanDecreaseAccuracy, mdg_community = MeanDecreaseGini),
  genus_supervised %>% importance() %>% as.data.frame() %>% rownames_to_column(var = "variables") %>% 
    select(MeanDecreaseAccuracy, MeanDecreaseGini,variables) %>% 
    mutate(rf_type = "genus_supervised") %>% 
    rename(mda_genus = MeanDecreaseAccuracy, mdg_genus = MeanDecreaseGini),
  by = "variables"
) %>% 
  write_csv("outputs/random_forest/RF_impotance_per_variables.csv")

```
The RF csv output has a missing value at genus supervised importance in the genus varibles because genus is the predicted classification variable.   

## Tidying the results output
```{r results_output}
plot_1_community_genus <- left_join(
  community_structure %>% importance() %>% as.data.frame() %>% rownames_to_column(var = "variables") %>% 
    select(MeanDecreaseAccuracy, MeanDecreaseGini,variables) %>% 
    mutate(rf_type = "community_structure") %>% 
    filter(variables != "genus") %>% 
    rename(mda_community = MeanDecreaseAccuracy, mdg_community = MeanDecreaseGini),
  genus_supervised %>% importance() %>% as.data.frame() %>% rownames_to_column(var = "variables") %>% 
    select(MeanDecreaseAccuracy, MeanDecreaseGini,variables) %>% 
    mutate(rf_type = "genus_supervised") %>% 
    rename(mda_genus = MeanDecreaseAccuracy, mdg_genus = MeanDecreaseGini, var_gen = variables),
  by = c("variables" = "var_gen")
) %>% 
  mutate(point_color = factor(case_when(
    .$variables %in% c("seq_method", "authors", "region", "growth_form", "health_status") ~ "Abiotic",
    str_detect(.$variables, "Candidat")                                                   ~ "Candidatus",
    str_detect(.$variables, "candidat")                                                   ~ "Candidatus",
    TRUE                                                                                  ~ "Established"
  ))) %>% 
  mutate(variables = factor(
    str_replace(.$variables, pattern = "\\.", replacement = " ") %>%
      str_replace(pattern = "_", replacement = " "))) %>% 
  mutate(variables = str_replace(variables, "Candidatus", "Ca."))
```



## Ploting results
```{r plot_result_rf, fig.width=25.19684, fig.height=12.59844}
rf_plot <- ggplot(plot_1_community_genus, aes(x = mda_genus, y = mda_community, colour = point_color))+
  #geom_abline(slope = 1, intercept = 0, alpha = 0.5, color = "gray85")+
  geom_point(size = 2.5)+
  ggrepel::geom_text_repel(data = filter(plot_1_community_genus,
                           mda_community > 50 | mda_community < 5 | mda_genus > 50 | point_color == "Abiotic" | variables %in% c("Cyanobacteria", "Alphaproteobacteria", "Betaproteobacteria", "Gammaproteobacteria", "Bacterioidetes")),force = 6, family = "sans", show.legend = FALSE,
             aes(x = mda_genus, y = mda_community, label = variables),
             size = 4)+
  labs(x = "Importance in separating coral genus",
       y = "Importance in community structure",
       colour = NULL)+
  coord_equal() +
  guides(colour = guide_legend(override.aes = list(size = 7.5))) +
  scale_colour_manual(values = c("Abiotic" = "#d95f02", "Candidatus" = "#60A260",
                                 "Established" = "#386CB0"))+
  #expand_limits(
  #  x = c(0, 73)
  #)+ 
  theme(axis.title.x = element_text(size=12, family = "sans"),
        panel.grid.major = element_blank(),panel.background = element_blank(),
        axis.line.y.left = element_line(), axis.line.x.bottom = element_line(),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 12, family = "sans"),
        axis.text.y = element_text(size = 10, family = "sans"),
        axis.text.x = element_text(size = 10, family = "sans"))+
  theme(axis.title.y = element_text(size=12, family = "sans"),
        #panel.grid.major.y = element_line(colour = "black")
        )

ggplot2::ggsave(plot = rf_plot, file="outputs/random_forest/Random_Forest_model_importance_community_and_genus.svg", width = 88, height = 45, units = "mm", scale = 4)
ggplot2::ggsave(plot = rf_plot, file="outputs/random_forest/Random_Forest_model_importance_community_and_genus.png", width = 88, height = 88, units = "mm", scale = 2, dpi = 250)
ggplot2::ggsave(plot = rf_plot, file="outputs/random_forest/Random_Forest_model_importance_community_and_genus.pdf", width = 88, height = 45, units = "mm", scale = 4)
rf_plot
```

```{r plot_result_abundance, fig.width=25.19684, fig.height=12.59844}

abund_media <- data_num %>% gather() %>% group_by(key) %>% 
  summarise(media = mean(value)) %>% 
  rename(variables = key)
write_csv(abund_media, "outputs/tables/mean_abun.csv")

df_abund_imp <- plot_1_community_genus %>% 
  filter(point_color != "Abiotic") %>% 
  select(mda_community, point_color, variables) %>% 
  arrange(variables) %>% 
  bind_cols(abund_media) %>% 
  mutate(variables = str_replace(variables...3, "Candidatus", "Ca."))
  
abund_comm <- df_abund_imp %>% 
  ggplot(aes(x = media, y = mda_community))+
  labs(x = "Average relative abundance (log10)",
       y = "Importance in community structure",
       colour = "Taxa")+
  #geom_hline(
  #  yintercept = df_abund_imp %>% mutate(rank_rel = rank(mda_community)) %>% 
  #    filter(rank_rel == 159%/%2) %>% pull(mda_community),
  #  linetype = 3
  #) +
  scale_x_log10(labels = trans_format("log10", math_format(10^.x)))+
  #annotation_logticks(colour = "gray80",sides = "bottom")+
  scale_colour_manual(values = c("Candidatus" = "#60A260",
                                 "Established" = "#386CB0"))+
  #guides(colour = "none") +
  geom_point(aes(colour = point_color), size = 2.5)+
  ggrepel::geom_text_repel(data = df_abund_imp %>% 
                             filter(mda_community >= 50| media >= 0.05| variables %in% c("Cyanobacteria", "Alphaproteobacteria", "Betaproteobacteria", "Gammaproteobacteria", "Bacterioidetes") | mda_community < 5), force = 15, show.legend = FALSE,
                           aes(colour = point_color, label = variables),
                           size = 4, family = "sans")+
  geom_vline(xintercept = .1/100, linetype = 3)+
  expand_limits(
    y = c(- 1, 90)
  )+
  theme(axis.title.x = element_text(size=12, family = "sans"),
        panel.grid.major = element_blank(),panel.background = element_blank(),
        axis.line.y.left = element_line(), axis.line.x.bottom = element_line(),
        panel.grid.minor = element_blank(), axis.ticks.length.x.top = element_blank(),
        legend.position = "bottom", legend.title = element_blank(),
        legend.text = element_text(size = 12, family = "sans"),
        axis.text.y = element_text(size = 10, family = "sans"),
        axis.text.x = element_text(size = 10, family = "sans"))+
  theme(axis.title.y = element_text(size=12, family = "sans"),
        #panel.grid.major.y = element_line(colour = "black")
        )

ggplot2::ggsave(abund_comm, file = "outputs/random_forest/mean_abundance_imp_community.png", width = 160, height = 80, units = "mm", scale = 4)
ggplot2::ggsave(abund_comm, file = "outputs/random_forest/mean_abundance_imp_community.pdf", width = 160, height = 80, units = "mm", scale = 4)
ggplot2::ggsave(abund_comm, file = "outputs/random_forest/mean_abundance_imp_community.svg", width = 160, height = 80, units = "mm", scale = 4)
abund_comm

```

## C plot

```{r plot C}
plot_c_df <- plot_1_community_genus %>% 
  filter(point_color != "Abiotic") %>% 
  #select(mda_community, point_color, variables) %>% 
  arrange(variables) %>% 
  left_join(
    abund_media %>% 
      mutate(point_color = factor(case_when(
    .$variables %in% c("seq_method", "authors", "region", "growth_form", "health_status") ~ "Abiotic",
    str_detect(.$variables, "Candidat")                                                   ~ "Candidatus",
    str_detect(.$variables, "candidat")                                                   ~ "Candidatus",
    TRUE                                                                                  ~ "Established"
  ))) %>% 
  mutate(variables = factor(
    str_replace(.$variables, pattern = "\\.", replacement = " ") %>%
      str_replace(pattern = "_", replacement = " "))) %>% 
  mutate(variables = str_replace(variables, "Candidatus", "Ca."))
            ) 

plot_c <- plot_c_df %>% 
  ggplot(aes(x = media, y = mda_genus))+
  labs(x = "Average relative abundance (log10)", 
       y = "Importance in separating coral genus",
       colour = "Taxa")+
  scale_x_log10(labels = trans_format("log10", math_format(10^.x)))+
  #annotation_logticks(colour = "gray80",sides = "bottom")+
  scale_colour_manual(values = c("Candidatus" = "#60A260",
                                 "Established" = "#386CB0"))+
  guides(colour = "none") +
  #geom_hline(
  #  yintercept = plot_c_df %>% mutate(rank_rel = rank(mda_genus)) %>% 
  #    filter(rank_rel == 159%/%2) %>% pull(mda_genus),
  #  linetype = 3
  #) +
  geom_point(aes(colour = point_color), size = 2.5)+
  ggrepel::geom_text_repel(data = plot_c_df %>% 
                             filter(mda_genus >= 50| media >= 0.05| variables %in% c("Cyanobacteria", "Alphaproteobacteria", "Betaproteobacteria", "Gammaproteobacteria", "Bacterioidetes") | mda_genus < 5), force = 15,
                           aes(colour = point_color, label = variables),
                           size = 4, family = "sans")+
  geom_vline(xintercept = .1/100, linetype = 3)+
  expand_limits(
    y = c(- 1, 90)
  )+
  theme(axis.title.x = element_text(size=12, family = "sans"),
        panel.grid.major = element_blank(),panel.background = element_blank(),
        axis.line.y.left = element_line(), axis.line.x.bottom = element_line(),
        panel.grid.minor = element_blank(), axis.ticks.length.x.top = element_blank(),
        legend.position = "bottom", legend.title = element_blank(),
        legend.text = element_text(size = 12, family = "sans"),
        axis.text.y = element_text(size = 10, family = "sans"),
        axis.text.x = element_text(size = 10, family = "sans"))+
  theme(axis.title.y = element_text(size = 12, family = "sans"),
        #panel.grid.major.y = element_line(colour = "black")
        )


ggplot2::ggsave(filename = paste0("outputs/random_forest/B_and_C_",Sys.Date(),".png"), height = 160, width = 80, units = "mm", scale = 4)
ggplot2::ggsave(filename = paste0("outputs/random_forest/B_and_C",Sys.Date(),".svg"), height = 160, width = 80, units = "mm", scale = 4)
ggplot2::ggsave(filename = paste0("outputs/random_forest/B_and_C",Sys.Date(),".pdf"), height = 160, width = 80, units = "mm", scale = 4)
```


## Final Panel
```{r Panel, fig.width=25.19684, fig.height=12.59844}

rf_plot / (abund_comm | plot_c) + 
  plot_annotation(tag_levels = "a", tag_prefix = "1") +
  plot_layout(guides = 'collect', nrow = 2) & theme(legend.position = 'bottom')
  

#plot_grid(rf_plot, abund_comm, plot_c, rel_widths = c(1,.5,.5), ncol = 1, 
#          #align = "h",
#          nrow = 2,labels = c("A", "B"),label_x = c(0,-0.01))

ggplot2::ggsave(filename = paste0("outputs/random_forest/rf_panel_",Sys.Date(),".png"), width = 160, height = 80, units = "mm", scale = 4)
ggplot2::ggsave(filename = paste0("outputs/random_forest/rf_panel_",Sys.Date(),".svg"), width = 160, height = 80, units = "mm", scale = 4)
ggplot2::ggsave(filename = paste0("outputs/random_forest/rf_panel_",Sys.Date(),".pdf"), width = 160, height = 80, units = "mm", scale = 4)
```




```{r fig 2, fig.height=25.19684, fig.width=12.59844}
fig_2 <- abund_comm + plot_c +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = 'collect', nrow = 1) & theme(legend.position = 'bottom', legend.box.background = element_blank())

ggplot2::ggsave(plot = fig_2, file="outputs/random_forest/importance_community_and_genus_abund.png", width = 180, height = 90, units = "mm", scale = 2, dpi = 250)
ggplot2::ggsave(plot = fig_2, file="outputs/random_forest/importance_community_and_genus_abund.svg", width = 180, height = 90, units = "mm", scale = 2, dpi = 250)
```


