---
title: "Holobiont project KW"
author: "Arthur & Leticia"
date: "12/10/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

# Kruskal-Wallis Univariate test
## Libraries we used
```{r libraries, warning=FALSE,message=FALSE}
library(janitor)
library(tidyverse)
library(gridExtra)
library(Matrix)
library(broom)
library(kableExtra)
```

## Importing and tidying the data
```{r reading data}
data <- read.csv("input/detailed_abundance_phyla_157_corais_matrix_porcentagem_prot_divided_17abr2020.csv",
                 check.names = FALSE) %>% 
  select(-1)

levels(data$health_status)[levels(data$health_status) == "iminent bleaching"] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "signs of bleaching"] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "bleaching event"] <- "bleached" 
levels(data$health_status)[levels(data$health_status) == "bleaching event "] <- "bleached"
levels(data$health_status)[levels(data$health_status) == "Cyanobacterial patches"] <- "Black Band Disease"


data_num <- data %>% select_if(is.numeric)

abundantes <- data_num %>% gather() %>% group_by(key) %>% 
  summarise(mean_abund = mean(value)) %>% 
  filter(mean_abund > 1/100) %>% pull(key)


data <- data %>% 
  mutate(
    seq_method = factor(
      case_when(
        .$seq_method %in% c("Illumina Genome Analyzer IIx", "Illumina HiSeq 1000",
                            "Illumina HiSeq 1000 WGS", "Illumina HiSeq 2000",
                            "Illumina HiSeq 2500", "Illumina HiSeq 2500 WGS",
                            "WGS illumina") ~ "Illumina",
        TRUE ~ "454"
      )
    )
  )

data <- data %>% filter(genus != "Fungia",
                        genus != "Pseudodiploria", 
                        !(.$genus == "Porites" & .$authors == "Thurber et al. 2009"),
                        !(.$genus == "Acropora" & .$region == "mesocosmos" & .$authors == "not published"),
                        !(.$genus == "Pocillopora" & .$region == "mesocosmos" & .$authors =="not published"),
                        !(.$genus == "Seriatopora" &.$region == "mesocosmos" & .$authors =="not published"),
                        !(.$id == "mgm4694757_prinseq_good_Crf1")) %>% 
  mutate(genus = factor(genus),
         authors = factor(authors),
         growth_form = factor(growth_form))

data_num <- data %>% select_if(is.numeric)
```

## Kruskal-Wallis Multivariate test Function
```{r mult_KW_function}
multkw <- function(group,y,simplify=FALSE){
  
  ### sort data by group ###
  o<-order(group)
  group<-group[o]
  y<-as.matrix(y[o,])
  n<-length(group)
  p<-dim(y)[2]
  if (dim(y)[1] != n)
    return("number of oberservations not equal to length of group")
  
  groupls<-unique(group)
  g<-length(groupls) #number of groups#
  groupind<-sapply(groupls,"==",group) #group indicator#
  ni<-colSums(groupind) #num of subj of each group#
  r<-apply(y,2,rank) #corresponding rank variable#
  
  ### calculation of statistic ###
  r.ik<-t(groupind)%*%r*(1/ni) #gxp, mean rank of kth variate in ith
  group#
  m<- (n+1)/2 #expected value of rik#
  u.ik<-t(r.ik-m)
  U<-as.vector(u.ik)
  V<-1/(n-1)*t(r-m)%*%(r-m) #pooled within-group cov matrix
  Vs<-bdiag(lapply(1/ni,"*",V))
  W2<-as.numeric(t(U)%*%solve(Vs)%*%U)
  
  ### return stat and p-value ###
  returnlist<-list(statistic=W2,d.f.=p*(g-1),
                   p.value=pchisq(W2,p*(g-1),lower.tail=F))
  if (simplify==TRUE) return (W2)
  else return (returnlist)
}
```

## PCA

- First we scale and center the abundance matrix
- Then we calculate the covariation matrix
- So we can extract the eigenvalues and eigenvectors from the covariation matrix

```{r manual approach}
data_scale <- scale(data_num, center = TRUE, scale = TRUE)
data_cov <- cov(data_scale)
#det(data_cov)

auto <- eigen(data_cov)

#var.exp <- cumsum(auto$values/sum(auto$values))
#sum(var.exp <= .9)

PCs <- as.matrix(data_num) %*% matrix(auto$vectors[,1:33],nrow=159,ncol=33)
```

- Then we filter the rincipal components that don't passs our **treshhold** 
  + Eigenvalues greather than 1
  
```{r verify eigen}
sum(auto$values>1)
```
- With the 33 PCs we perform the univariate and the multivariate kruskal-wallis tests

### Multivariate Kruskal-Wallis
```{r MKW}
mkw_aut <- multkw(y = PCs, group = data$authors)
mkw_genus <- multkw(y = PCs, group = data$genus)
mkw_health <- multkw(y = PCs, group = data$health_status)
mkw_seq <- multkw(y = PCs, group = data$seq_method)
mkw_GF <- multkw(y = PCs, group = data$growth_form)
mkw_region <- multkw(y = PCs, group = data$region)

mkw_geral <- bind_rows(mkw_aut, mkw_genus,
                       mkw_health,mkw_seq,
                       mkw_GF, mkw_region)

mkw_geral$var <- c("authors", "genus", "health_status",
                   "seq_method", "growth_form", "region")

mkw_geral %>% 
  arrange(desc(statistic)) %>% 
  kbl(caption = "MKW test results",digits = 1e100) %>%
  kable_classic(full_width = F, html_font = "Cambria")
mkw_geral %>% write_csv("outputs/table_mkw_geral.csv")

```

### Univariate Kruskal-Wallis 

```{r individual test}
pvalue <- matrix(0, nrow=ncol(PCs), ncol=length(2:8)) %>% as.data.frame()
statistic <- pvalue
d_f <- pvalue

for(i in 1:ncol(PCs)){
  
  for(k in 2:8){
    
    le_test <- kruskal.test(PCs[, i] ~ data[, k])
    
    pvalue[i, k] <- le_test$p.value
    statistic[i, k] <- le_test$statistic
    d_f[i, k] <- le_test$parameter 
  }
  
}

df_kw_individual <- bind_cols(
  pvalue %>% rename_with( ~ paste("pvalue", .x, sep = "_")),
  statistic %>% rename_with( ~ paste("statistic", .x, sep = "_")),
  d_f %>% rename_with( ~ paste("dof", .x, sep = "_")),
) %>% mutate(PC = paste0("PC_", 1:n())) %>% 
  select(-c(pvalue_V1, statistic_V1, dof_V1))

result_kw <- df_kw_individual %>% select(contains("pvalue"), PC) %>% 
  pivot_longer(-PC) %>% 
  mutate(th_sig = value <= .05,
         variable = case_when(
           name == "pvalue_V2" ~ "genus",
           name == "pvalue_V3" ~ "species_host",
           name == "pvalue_V4" ~ "health_status",
           name == "pvalue_V5" ~ "growth_form",
           name == "pvalue_V6" ~ "seq_method",
           name == "pvalue_V7" ~ "region",
           TRUE ~ "authors"
         )
  ) %>% 
  group_by(variable) %>% count(th_sig) %>% 
  mutate(prop = n/ncol(PCs)) %>% filter(th_sig == TRUE) %>% 
  arrange(desc(prop)) %>% 
  select(variable, n, prop)

result_kw %>% 
  kbl(caption = "") %>% 
  kable_classic(full_width = F, html_font = "Cambria")

write_csv(x = result_kw, "outputs/table_kw_univar.csv")
```
table. Proportion of PCs that passed the arbitrary treashhold of 0.05 in the Kruskal-Wallis test
